---
title: "haps"
author: "Yasmin Hawar Abo Bakir Shuan, Eiríkur Jonsson & Camilla Dyhr-Jahns"
date: "18/5/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(st514)
library(tidyverse)
library(ggplot2)
library(GGally)
library(corrplot)
library(ggcorrplot)
library(ggpubr)
library(lattice)
data("T1-6")
df <- tbl
# Non-multiple-sclerosis (NMS) group data
nms <- subset(df, V6 == 0)
# Multiple-sclerosis (MS) group data
ms <- subset(df, V6 == 1)
```

```{r}
colnames(tbl) <- c("Age", "V2", "V3", "V4", "V5","target")
tbl$target <- as.factor(tbl$target)
```

Data related to responses to visual stimuli were used to develop a rule for seperating people suffering from a multiple-sclerosis-caused visual pathology from those not suffering from the disease.

We were unable to find a data description, so we assume that this dataset is measuring response time in milliseconds between the two samples, where V2 is a measurement for stimulus 1 in milliseconds and V3 is the difference in respeonse time for each eye. This is also the case for V4 and V5, respectively and V6 our label (0 for Non-MS and 1 for MS).

# DESCRIPTIVE STATISTICS


```{r include=FALSE}
descriptive <- function(x){
  xmean <- colMeans(x)
  xcov <- cov(x)
  xcovVar <- cov(x) * (nrow(x) -1) / nrow(x)
  xvar <- var(x)
  xcovInv <- solve(xcov)
  xdet <- det(xcov)

  # Debug block
  
  # print("Mean Vector")
  # print(xmean)
  # print("-----------------------------------------------------")
  # print("Covariance")
  # print(xcov)
  # print("-----------------------------------------------------")
  # print("Xco/Variance")
  # print(xcovVar)
  # print("-----------------------------------------------------")
  # print("Variance")
  # print(xvar)
  # print("-----------------------------------------------------")
  # print("Covariance inverse")
  # print(xcovInv)
  # print("-----------------------------------------------------")
  # print("Determinant")
  # print(xdet)
  mylist = list("meanVector" = xmean, "Covariance" = xcov, "Co-Variance" = xcovVar, "Variance" = xvar, "CovInverse" = xcovInv, "Determinant" = xdet)
  return(mylist)
  
}
```

## FOR NMS GROUP
The mean age for the NMS group is lower than for the MS group.
The responses to stimuli 1 and 2 are faster in the NMS group compared to the MS group.
Finally, the differences between L and R are significantly higher in the MS group compared to the NMS group.

### COVARIANCE
V2 and V4
The covariance for the NMS sample is 106.78 whereas for the MS sample it is 912.41, which indicates a slower responsetime in the MS group. 

V3 and V5
When calculating the difference in respeonse time for each eye, the NMS group has differnce of 0.50. The MS group has a difference of 297.31 which again indicates a high difference between the two groups.

### VARIANCE-COVARIANCE

A variance-covariance matrix is a square matrix that contains the variances and covariances associated with several variables. The diagonal elements of the matrix contain the variances of the variables and the off-diagonal elements contain the covariances between all possible pairs of variables.
The variance-covariances associated with variables V2 and V4 for the NMS group is 105.23. 
MS
880.95234

V3 and V5
NMS
0.4940979
MS
287.06683

### VARIANCE
Variance tells us how much a quantity varies w.r.t. its mean. Its the spread of data around the mean value. You only know the magnitude here, as in how much the data is spread.
V2 and V4
NMS
106.785030
MS 
V3 and V5


```{r}
descriptive(nms[1:5])
```
```{r}
descriptive(ms[1:5])
```
```{r}
print(var(ms[1:5]))
print(cov(ms[1:5]))
```

```{r}
Normal_Distributed <- hist(nms$V2+nms$V4)

# The varibles, V2 and V4 from the Non-MS group has the highest correlation. The plot below shows that they are normally distributed.
```


```{r}
library(devtools)
install_version('MVN', '4.0.2')
```


```{r}
qqnorm(nms[, 2])
qqline(nms[, 2])
```


```{r}
# (Note: we have plottet all the variables, V2 and V4, and V3 and V5, from both groups and only the V2 and V4 from the Non-MS group have normally distributed data and therefore the only one shown).
```





