---
title: "multiExam"
author: "Eirikur Jonsson"
date: "09/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(st514)
library(tidyverse)
library(ggplot2)
library(GGally)
library(corrplot)
library(ggcorrplot)
library(MASS)
library(rcompanion)
library(cowplot)
library(psych)
library(boot)
library(caret)
source("https://imada.sdu.dk/~chdj/ST514/F20/BoxM.r")
data("T1-6")
df <- tbl
# Non-multiple-sclerosis (NMS) group data
nms <- subset(df, V6 == 0)
# Multiple-sclerosis (MS) group data
ms <- subset(df, V6 == 1)
```

Add Descriptive function here

# Data Preperation

## Testing for normality with qqplots Mahalanobis SD whole and half/ass

```{r}
x <- as.matrix(tbl[1:5])
x.mean <- colMeans(x)
S <- cov(x)
n <- nrow(x)
D2 <- rep(0, n)
for (i in 1:n) {
  D2[i] <- t(x[i,] - x.mean) %*% solve(S) %*% (x[i,] - x.mean)
}
#round(D2, 4)
#which(D2 < qchisq(0.5, 2))
p <- ncol(x)
D2 <- mahalanobis(x, colMeans(x), S)
result <- qqplot(qchisq(ppoints(n, a = 0.5), df = p), D2,
       ylab = "Mahalanobis distances",
       xlab = bquote("Quantiles of " ~ chi[.(p)]^2),
       main = bquote("Q-Q plot of Mahalanobis" * ~ D^2 * 
                       " vs. quantiles of" * ~ chi[.(p)]^2))
cor(result$x, result$y)
```


### Are the subsets means possible mean vectors in each.

Using Hotellings $T^2$

Here we test these hypotheses:

$$
h_0:\mu(MS) = \mu(NMS) \\
h_1: \mu(MS) \ne \mu(NMS)
$$

$$
h_0: \mu(NMS) = \mu(MS) \\
h_1: \mu(NMS) \ne \mu(MS)
$$

```{r}
msMatrix <- as.matrix(ms[1:5])
nmsMatrix <- as.matrix(nms[1:5])
msMeans <- colMeans(msMatrix)
nmsMeans <- colMeans(nmsMatrix)
msN <- nrow(msMatrix)
nmsN <- nrow(nmsMatrix)
msP <- ncol(msMatrix)
nmsP <- ncol(nmsMatrix)
msCov <- cov(msMatrix)
nmsCov <- cov(nmsMatrix)


# Start -  are ms mu possible mean vectors in nms
msInNms <- t(nmsMeans - msMeans)%*%solve(nmsCov)%*%(nmsMeans - msMeans)
msInNmsT2 <- round(nmsN * msInNms, 3)
msInNmsFval <- msInNmsT2 * (nmsN - nmsP)/((nmsN - 1)*nmsP)
msInNmsFvalP <- pf(msInNmsFval, nmsP, nmsN - nmsP, lower.tail = FALSE)
# Now we do the reverse
nmsInMs <- t(msMeans - nmsMeans)%*%solve(msCov)%*%(msMeans - nmsMeans)
nmsInMsT2 <- round(msN * nmsInMs, 3)
nmsInMsFval <- nmsInMsT2 * (msN - msP) / ((msN - 1) * msP)
nmsInMsFvalP <- pf(nmsInMsFval, msP, msN - msP, lower.tail = FALSE)

print(paste("Nms in MS ", msInNmsT2, "With p value ", msInNmsFvalP))
print(paste("MS in NMS ", nmsInMsT2, "with a p value ", nmsInMsFvalP))
```

From these we can see that we can reject the null hypotheses in both instances that the means from each subset is a possible mean vector in the other.

## Outliers
Using boxplots to see if there are outliers
```{r}
df <- tbl
colnames(df) <- c(paste("x", 1:5, sep=""), "d2")
dfz <- round(scale(df[1:5]), 1)
colnames(dfz) <- c(paste("z", 1:5, sep = ""))
outliers <- cbind(df[,1:5], 1:98, dfz, tbl[,5])
colnames(outliers)[c(6,12)] <- c("Obs", "d2")
max(outliers$d2)
which(outliers$d2 == max(outliers$d2))

#This function calculates the critical value
chi_crit_val <- qchisq(0.005, 5, lower.tail = F)
chi_crit_val

#This function prints out the outliers by observation number in relation to the treshold of 16.74
out <- which(outliers$d2 > chi_crit_val)
which(outliers$d2 > 16.74)
df <- tbl[-out,]
df1 <- scale(tbl[1:5])
df1 <- cbind(df1, tbl$V6)
```

Function to standardize the values of each column

```{r}
colnames(df1)[6] <- c("V6")
head(df1)
```

```{r}
df <- as.data.frame(df1)
colnames(df) <- c("x1", "x2","x3", "x4", "x5", "d2")
pca <- prcomp(~ x1 + x2 + x3 + x4 + x5, data = df, scale = TRUE)

loadings <- pca$rotation

axes <- predict(pca, newdata = df)
```
```{r}
dat <- cbind(df, axes)
dat$d2 <- as.factor(dat$d2)
train <- createDataPartition(dat$d2, p = 0.6, list = FALSE)
training <- dat[train,]
testing <- dat[-train,]

model_log <- train(d2~ PC1 + PC2, data = training, method = "glm", family = "binomial")
model_log

pred <- predict(model_log, newdata = testing)
caret::confusionMatrix(data = pred, testing$d2)
```


```{r}
qda_model <- qda(d2~PC1+PC2,data = dat2, CV = TRUE)
a <- table(dat2$d2, qda_model$class)
print("Confusion Matrix")
print(a)
print(paste("Accuracy: ", sum(diag(a))/nrow(dat2)))
```


```{r}
BoxM(dat[1:5], dat[6], Ftest = FALSE)
```